Higher order unification for free!

### 2024-04-09 paper/main.tex
rework of a proof

\newcommand{\progressetadedup}{\emph{progress-$\eta$-deduplicate}}
\begin{definition}[\progressetadedup]\label{def:progressetadedup}
  A link \linketaM{\Gamma}{X_{\vec{s}}}{T} is removed from \linkStore when
  another link \linketaM{\Delta}{X_{\vec{r}}}{T'} is in  \linkStore.
  By \cref{inv:uvaarity} the length of $\vec{s}$ and $\vec{r}$ is the same
  hence we can move the term $T'$ from $\Delta$ to $\Gamma$ by renaming its
  bound variables, i.e. $T'' = T'[\vec{r}/\vec{s}]$.
  We then run $T \Ue{} T''$ (under the context $\Gamma$).
\end{definition}

@@

\begin{proof}[Proof sketch]
  By \cref{lemma:unif-eta-1,lemma:unif-eta-2,lemma:unif-eta-3}, every
  unification performed by the activation of a \linketa is performed between
  terms in \wellb, therefore, the substitution remains \wellb.
  \noteD[inline]{Bisogna aggiungere un lemma nella \cref{sec:nutshell} che dice che
  unificare due termini in \wellb, in una $\sigma$, tale che $\wellb(\sigma)$, 
  non invalida \wellb}
\end{proof}

@@

\noteD[inline]{Proove simulation fidelity, dicendo che \progressetaright è inutile}

\begin{lemma}[Fidelity with \linketa]
  Let \hoUnifPb a unification problem and $\sigma$ a substitution. The introduction 
  of \linketa guarantees \cref{prop:fidelity} 
  \label{lemma:fidelity-maybeeta}
\end{lemma}

\begin{proof}[Proof sketch]
  If 1) a \linketa is activated by \progressetaleft, then unification is done
  between a \maybeeta term $t_1$ and a term $t_2$ (with $\wellb(t_2)$). This
  activation performs a unification which succeeds iff the original problem in
  \Fo{} succeeds. If 2) a \linketa is activated by \progressetadedup, then the
  unification is done between two \maybeeta terms, and again this unification
  succeeds iff it succeeds in \Fo{}. Finally, if 3) a \linketa is activated by
  \progressetaright, the unification, done between a variable and a term,
  always succeeds, this is what we expect to guarantee fidelity, since we
  are essentially removing a hole added by the compilation in the place of a
  \maybeeta subterm. In all the cases fidelity is respected.
\end{proof}

>>>>

\newcommand{\progressetadedup}{\emph{progress-$\eta$-deduplicate}\xspace}
\begin{definition}[\progressetadedup]\label{def:progressetadedup}
  A link \linketaM{\Gamma}{X_{\vec{s}}}{T} is removed from \linkStore when
  another link \linketaM{\Delta}{X_{\vec{r}}}{T'} is in  \linkStore.
  By \cref{inv:uvaarity} the length of $\vec{s}$ and $\vec{r}$ is the same
  hence we can move the term $T'$ from $\Delta$ to $\Gamma$ by renaming its
  bound variables, i.e. $T'' = T'[\vec{r}/\vec{s}]$.
  We then run $T \Ue{} T''$ (under the context $\Gamma$).
\end{definition}

@@

\begin{proof}[Proof sketch]
  By \cref{lemma:unif-eta-1,lemma:unif-eta-2,lemma:unif-eta-3}, every
  unification performed by the activation of a \linketa is done between
  terms in \wellb, therefore, the substitution remains \wellb.
  \noteD[inline]{Bisogna aggiungere un lemma nella \cref{sec:nutshell} che dice che
  unificare due termini in \wellb, in una $\sigma$, tale che $\wellb(\sigma)$, 
  non invalida \wellb}
\end{proof}

@@

\begin{lemma}[Fidelity with \linketa]
  The introduction 
  of \linketa guarantees \fullref{prop:fidelity} 
  \label{lemma:fidelity-maybeeta}
\end{lemma}

\begin{proof}[Proof sketch]
  \progressetaleft and \progressetadedup activate a \linketa when, in the
  original unification problem, a \maybeeta term is unified with respectively a
  well-behaved term or another \maybeeta term. In both cases, the links trigger
  a unification which succeeds iff the same unification in \Fo{} succeeds,
  guaranteeing \cref{prop:fidelity}. \progressetaright never fails, in fact,
  this progression refines a variable to a rigid term and plays no role in 
  \cref{prop:fidelity}.
  % If 1) a \linketa is activated by \progressetaleft, then unification is done
  % between a \maybeeta term $t_1$ and a term $t_2$ (with $\wellb(t_2)$). This
  % activation performs a unification which succeeds iff the original problem in
  % \Fo{} succeeds. If 2) a \linketa is activated by \progressetadedup, then the
  % unification is done between two \maybeeta terms, and again this unification
  % succeeds iff it succeeds in \Fo{}. Finally, if 3) a \linketa is activated by
  % \progressetaright, the unification, done between a variable and a term,
  % always succeeds, this is what we expect to guarantee fidelity, since we
  % are essentially removing a hole added by the compilation in the place of a
  % \maybeeta subterm. In all the cases fidelity is respected.
\end{proof}
